---
title: "Project 2: Transformation"
author: "Jordan Tapke"
date: "9/27/2020"
output:
  html_document: 
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Set 1: Which Streaming Platform Has the Best Movies?

### Overview

The goal for this data set was to tidy up the data and analyze which streaming provider has better movies on average.

### Import and Subet Data to Needed Columns
```{r, message=FALSE}
#import data from github
movies_df <- read.csv(url('https://raw.githubusercontent.com/jtapke/School-Projects/master/Project%202/Movies%20On%20Streaming%20Platorms/MoviesOnStreamingPlatforms_updated.csv'))

#load required libraries
library(tidyverse)

#subset data for analysis
myvars <- c(2,6:11)
movies_df <- movies_df[myvars]
head(movies_df)
```

### Standardize and Combine Ratings
```{r}
#bring Rotten Tomato rating to 1-10 scale
rotten_rating <- as.numeric(sub("%","", movies_df$Rotten.Tomatoes))/10
imdb_rating <- movies_df$IMDb

#combine the the ratings from both providers into an overall rating
overall_rating <- (rotten_rating + imdb_rating)/2

#append overal rating to dataframe
movies_df <- movies_df %>%
  mutate(overall_rating, .after="ID" )

#remove old ratings
movies_df <- movies_df[c(1:2,5:8)]

head(movies_df)
```

### Separate Movies by Provider
```{r}
#separate movies by provider
netflix <- movies_df %>%
  filter(Netflix == 1)

hulu <- movies_df %>%
  filter(Hulu == 1)

prime <- movies_df %>%
  filter(Prime.Video == 1)

disney <- movies_df %>%
  filter(Disney. == 1)
```

### Aggregate Avg. Rating of All Movies on Each Platform
```{r}

netflix_avg <- mean(netflix$overall_rating, na.rm = TRUE)

hulu_avg <- mean(hulu$overall_rating, na.rm = TRUE)

prime_avg <- mean(prime$overall_rating, na.rm = TRUE)

disney_avg <- mean(disney$overall_rating, na.rm = TRUE)

#create data frame of avg rating of each platform
platform_ratings <- data.frame(netflix_avg,hulu_avg, prime_avg, disney_avg)

#change from wide to long format for analysis
platform_ratings <- platform_ratings %>%
  pivot_longer(netflix_avg:disney_avg, names_to = "platform", values_to = "avg_movie_rating")
```

### Barplot of Ratings
```{r}
ggplot(data=platform_ratings, aes(x=reorder(platform, -avg_movie_rating), y=avg_movie_rating, fill=platform)) +
  
  #allows for Y value to be variable instead of "count"
    geom_bar(stat="identity")+
  #changes labels of axis
    xlab("Platform") + ylab("Avg. Rating of Movies") +
  #creates title for graph
    ggtitle("Which Platform has the Best Movies?")+
  #changes labels on x values
    scale_x_discrete(labels=c("Netflix","Hulu","Dinsey+","Prime Video"))+
  #removes legend
    guides(fill=FALSE)+
  #zooms in on data points to better see if there are differences
    coord_cartesian(ylim = c(5,6.7))
```

### Summary 

Based on the data provided, Netflix has the highest average movie ratings on its platform while Prime Video has the lowest average movie ratings on its platform. However the differences are barely distinguishable if the graph was not zoomed in.


# Data Set 2: World Happiness Report - Which Factor Contributes most to Happiness?

### Overview

The goal for this data set was to combine and tidy up the data to analyze which factor in the world happiness report contributes most towards a country's happiness score on average.

### Import CSV Files
```{r}

for(i in c(2015:2019)) { 
 #assigning name to dataframe 
  assign(paste0("data_", i),
         
         #Reading csv file from github
         read.csv(url(paste0("https://raw.githubusercontent.com/jtapke/School-Projects/master/Project%202/World%20Happiness%20Report/", i, ".csv"))))
}
```


### Subset Columns and Rename
```{r}
#each file has different columns and orders
data_2015 <- data_2015[,c(4,6,9:11)]
colnames(data_2015) <-c( "total", "gdp", "freedom", "corruption", "generosity")

data_2016 <- data_2016[,c(4,7,10:12)]
colnames(data_2016) <-c( "total", "gdp", "freedom", "corruption", "generosity")

data_2017 <- data_2017[,c(3,6,9,11,10)]
colnames(data_2017) <-c( "total", "gdp", "freedom", "corruption", "generosity")

data_2018 <- data_2018[,c(3:4,7,9,8)]
colnames(data_2018) <-c( "total", "gdp", "freedom", "corruption", "generosity")

data_2019 <- data_2019[,c(3:4,7,9,8)]
colnames(data_2019) <-c( "total", "gdp", "freedom", "corruption", "generosity")
```

### Convert Data Types
```{r, warning=FALSE}
#convert to numeric so dataframes can be combined
data_2015[, 2:ncol(data_2015)] <- sapply(data_2015[, 2:ncol(data_2015)], as.numeric)
data_2016[, 2:ncol(data_2016)] <- sapply(data_2016[, 2:ncol(data_2016)], as.numeric)
data_2017[, 2:ncol(data_2017)] <- sapply(data_2017[, 2:ncol(data_2017)], as.numeric)
data_2018[, 2:ncol(data_2018)] <- sapply(data_2018[, 2:ncol(data_2018)], as.numeric)
data_2019[, 2:ncol(data_2019)] <- sapply(data_2019[, 2:ncol(data_2019)], as.numeric)
```

### Combine into One Dataframe
```{r}
library(tidyverse)

#new data frame
happiness_report <- data.frame(bind_rows(data_2015,data_2016,data_2017, data_2018, data_2019))
```

### Calculate Factor's Ratio of Total Score
```{r, message=FALSE, warning=FALSE}
library(data.table)
#calculate sum of each factor
total <- sum(happiness_report$total)
gdp <- sum(happiness_report$gdp)
freedom <- sum(happiness_report$freedom)
corruption <- sum(happiness_report$corruption, na.rm = TRUE)
generosity <- sum(happiness_report$generosity)
#create data frame
happiness_contributors <- data.frame(total,gdp,freedom, corruption, generosity)
#calculate ratio of total score
happiness_contributors <- happiness_contributors %>%
  mutate_at(vars(-total), funs(. / total)) %>%
  select(gdp, freedom, corruption, generosity)
```

### Pivot into Long Format Data Frame
```{r}

happiness_contributors_long <- happiness_contributors %>%
  pivot_longer(gdp:generosity, names_to = "factor", values_to = "percent")
happiness_contributors_long
```

### Barplot of Contributors to Happiness
```{r, message=FALSE}
require(scales)
ggplot(data=happiness_contributors_long, aes(x=reorder(factor, -percent), y=percent, fill=factor)) +
    geom_bar(stat="identity")+
    xlab("Factor") + ylab("Percent") +
    ggtitle("Average Percent Contributed to Happiness Score")+
    scale_x_discrete(labels=c("GDP","Freedom","Absence of Corruption","Generosity"))+
    scale_y_continuous(labels=percent_format())+
    guides(fill=FALSE)
```

### Summary 

Based on the data provided, GDP contributes the most towards a country's happiness score on average for all observations from 2016-2019. It contributes on average ~19% towards the happiness score. This is almost double than the second highest contributor, freedom, which was around ~7.5% of the happiness score.


# Dataset 3: Broadband Availibility in New York State

### Overview
The goal of this data was to find out which municipalities have lower rates of broadband and the number of housing units affected. However, I though this did not provide much insight so I also imported another data set of median income by county level to see if there was a correlation between a counties median income and its access to broadband internet.

### Import and Subet Data to Only Needed Columns
```{r, message=FALSE}
#load required libraries
library(tidyverse)

#import data from github
broadband_df <- read.csv(url('https://raw.githubusercontent.com/jtapke/School-Projects/master/Project%202/Broadband%20Availability%20By%20Municipality/Broadband_Availability_By_Municipality.csv'))

#add-on data set to bring more meaning to the original data set
income_county_df <- read.csv(url('https://raw.githubusercontent.com/jtapke/School-Projects/master/Project%202/Broadband%20Availability%20By%20Municipality/Income_ny_2014.csv'))

#subset broadband data for analysis
myvars <- c(7,2,4,20)
broadband_df <- broadband_df[,myvars]
colnames(broadband_df)<- c("county", "town", "population", "broadband_rate")

#subset income data for analysis
myvars <- c(2,246)
income_county_df <- income_county_df[,myvars]
income_county_df <- tail(income_county_df, -1)
income_county_df <- head(income_county_df, -1)


# tidy up county column
income_county_df <- extract(income_county_df, "NAME", into = c("county"),  regex = "(.*)\\sCounty.*")
colnames(income_county_df)<- c("county", "median_income")
```

### Total Population for Each County
```{r, message=FALSE}
#total population for each county
total_pop_df <- broadband_df %>%
  group_by(county) %>%
  summarise(pop = sum(population))

#remove unwanted variables
remove.list <- paste(c(",", "n/a"), collapse = "|")
total_pop_df <- total_pop_df %>%
  filter(!str_detect(total_pop_df$county, remove.list))
```
### People With Broadband Access per Town
```{r}
#create new variable: access_per_town
access_per_town <- #create new variable: access_per_town
access_per_town <- broadband_df$population * (broadband_df$broadband_rate/100)
broadband_df <- cbind(broadband_df, access_per_town )
```
### Broadband Access Rate per County
```{r, message=FALSE}
#total access for each county
access_per_county <- broadband_df %>%
  group_by(county) %>%
  summarise(access = sum(access_per_town))

#remove unwanted variables
remove.list <- paste(c(",", "n/a"), collapse = "|")
access_per_county <- access_per_county %>%
  filter(!str_detect(access_per_county$county, remove.list))

access_per_county <- cbind(access_per_county, total_pop_df$pop)

#create new variable: broadband_rate
broadband_rate <- access_per_county$access / access_per_county$total_pop
access_per_county <- cbind(access_per_county, broadband_rate )
```

### Join The Two Data Frames
```{r}
#join county data columns
county_inc_bb <- left_join(income_county_df, access_per_county, by = "county" )

#clean up data frame
median_inc_num <- as.numeric(county_inc_bb$median_income)
county_inc_bb <- cbind(county_inc_bb, median_inc_num)
county_inc_bb <- county_inc_bb[,c(1,5,6)]
```

### Scatterplot
```{r, message=FALSE}
ggplot(county_inc_bb, aes(median_inc_num, broadband_rate))+
  geom_point(color='steelblue', size=2)+
  geom_smooth(method = "loess", se = TRUE, color="brown3")+
  xlab("County Median Income")+ylab("Rate of Access")+ 
  ggtitle("Counties Rate of Broadband  Internet Access vs. Median Income")+
  scale_x_continuous(labels=scales::dollar_format())+
  scale_y_continuous(labels=scales::percent_format())
```

### Number of People with Broadband Internet
```{r}
popwo_access <- as.numeric(summarise(broadband_df, sum(population*(1-(broadband_rate/100)))))

popwo_access
```

### Summary 

Based on the data, there seems to be very little correlation between a county's median income and it's rate of access to broadband. What is still surprising is there there are over 3 million people in NYS who do not have access to broadband internet.


# References

Bhatia, R. (2020, May 22). Movies on Netflix, Prime Video, Hulu and Disney+. Retrieved October 03, 2020, from https://www.kaggle.com/ruchi798/movies-on-netflix-prime-video-hulu-and-disney

Network, S. (2019, November 27). World Happiness Report. Retrieved October 03, 2020, from https://www.kaggle.com/unsdsn/world-happiness

O. (2014, December 29). Broadband Availability By Municipality. Retrieved October 03, 2020, from https://data.ny.gov/Economic-Development/Broadband-Availability-By-Municipality/sjc6-ftj4

U. (2019). SELECTED ECONOMIC CHARACTERISTICS. Retrieved October 03, 2020, from https://data.census.gov/cedsci/table?q=ACSDP1Y2019.DP03+New+York
