---
title: "Assignment 6 - Web APIs"
author: "Jordan Tapke"
date: "10/21/2020"
output:
  html_document: 
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview
 The goal of this assignment was to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R data frame.

I used the GET function from  *httr* to read in the JSON data from the NYT API. I then used *RJSONIO* and *tidyverse* packages to help transform the data from JSON list into a tidy R data frame. 

### Load Needed Libraries
```{r, message=FALSE}
library(tidyverse)
library(RJSONIO)
library(httr)
```

### Read in JSON Data Using API Key
```{r}
res <- GET("https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json?api-key=rPZYH1OQvEnkwrDWG0VcKTqr8ENYHtIz")
```

### Transform JSON into R Data Frame
```{r}
data <- fromJSON(rawToChar(res$content))

data <- data$results

data <- data$books

json_df <- enframe(unlist(data))

books_df <- json_df %>%
  group_by(name) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = name, values_from = value) %>%
  select(-row)

books_df <- books_df[1:15,]

books_df
```

### Conclusion
It was important to convert the JSON data into a character list first using *rawToChar* so that I could use *fromJSON* function to de-serialize it into R objects. doing this allowed me to convert the data into a tidy R data frame.
