---
title: 'Assignment 2: SQL & R'
author: "Jordan Tapke"
date: "9/3/2020"
output:
  html_document: 
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Introduction

The Assignment requirements included:

  1. Survey friends on 6 movie titles and ask them to rank the titles from 1 to 5
  2. Import the collected data into R to store in a relational database
  3. Import the data from database into R for analysis

### Survey

To collect the data on movies I created a survey using [Google forms.](https://forms.gle/spNTAcBCkJv5zKFN8)

Survey results are available on [Github](https://github.com/jtapke/School-Projects/blob/master/MovieSurveyResponses.csv)

### Store Data in MySQL Database

I took these steps to upload the data into a MySQL database in Google Cloud:

1. Create connection to the database without saving password.
2. Create a table in the database for the data to be inserted in.
3. Import the data into a dataframe from the CSV file and amend the column names.
2. Write the dataframe to the database.

```{R Connection}
library(RMariaDB)

#Create Connection to database without showing password
con <- dbConnect(RMariaDB::MariaDB(),
                 dbname = 'movie_preferences',
                 user = 'jtapke', 
                 password = rstudioapi::askForPassword("Database password"), 
                 host = '34.75.89.1')
```

OPTION 1: CREATE TABLE IN DATABASE
```{r results='hide'}
#Create Table in database
dbSendQuery(con, "
CREATE TABLE IF NOT EXISTS movie_preferences.moviesurvey (
survey_id INT,
time CHAR(100),
parasite INT,
joker INT,
avengers_end_game INT,
crazy_rich_asians INT,
black_panther INT,
deadpool INT
);
")
```
OPTION 2: CREATE NORMALIZED SET OF TABLES IN DATABASE
```{r results='hide'}
#Create Tables in database
dbSendQuery(con, "
CREATE TABLE IF NOT EXISTS movie_preferences.surveyinfo (
survey_id INT NOT NULL AUTO_INCREMENT,
time CHAR(100),
PRIMARY KEY (survey_id)
);
")

dbSendQuery(con, "
CREATE TABLE IF NOT EXISTS movie_preferences.movies (
movie_id INT NOT NULL AUTO_INCREMENT,
movie CHAR(100),
PRIMARY KEY (movie_id)
);
")

dbSendQuery(con, "
CREATE TABLE IF NOT EXISTS movie_preferences.ratings (
survey_id INT,
movie_id INT,
rating INT,
FOREIGN KEY (survey_id) REFERENCES surveyinfo(survey_id),
FOREIGN KEY (movie_id) REFERENCES movies(movie_id)
);
")
```

```{R Importing Data}
#Create dataframe from csv file
moviesurvey <- read.csv(url("https://raw.githubusercontent.com/jtapke/School-Projects/master/MovieSurveyResponses.csv"))

#Change column names to match those in the database
colnames(moviesurvey) <- c("survey_id","time", "parasite","joker","avengers_end_game","crazy_rich_asians","black_panther","deadpool")

head(moviesurvey)
```

```{R Writing Data to Table}
#Write the dataframe to the database table created
dbWriteTable(con, name = "moviesurvey", value = moviesurvey, append = TRUE)
```

### Importing and Preparing Data

I took these steps to import and prepare the data for analysis:

1. Import table from database into dataframe.
2. Tidy up the data so there are now "movie" and "rating" columns while dropping the null values.

```{R Clean Up Data}
#Import table from database into dataframe
library(tidyr)
movieratings <- dbGetQuery(con, "SELECT * FROM moviesurvey")

long_movieratings <- movieratings%>% pivot_longer(3:8, names_to = 'movie', values_to = 'rating', values_drop_na = TRUE)
long_movieratings
```


### Analysis
Below allows us to visualize the data in a grouped bar plot. 


```{r histogram of ratings}
library(tidyverse)

ggplot(long_movieratings, aes(movie, rating), group = rating) +
  geom_bar(aes(fill=factor(rating)), position = "dodge", stat = "identity", width=.9)+
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_discrete(name = "Rating")


```

### Recomending Movies

One way we can use this data to recommend movie titles to the survey participants is to add a genre column to the movie table in the database. We can then match people who highly rated one genre to movie titles of the same genre. For Example: Survey participate #1 rated "Black Panther" as 5. Since this is a part of the superhero genre, we can recommend them another superhero movie that they have not watched such as "Joker."

### Conclusion

After completing this project I would recommend using scenario #2: creating a normalized database. This allows for more information to be added, such as a genre column, to the database without being redundant. 