---
title: "Assignment 3: Working with REGEXP"
author: "Jordan Tapke"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## #1

Scrape table from fivethirtyeight site
```{r table-capture}
#load rvest for webscraping
library(rvest)

url <- "https://projects.fivethirtyeight.com/mid-levels/college-majors/index.html?v=3&initialWidth=575&childId=framediv&parentTitle=The%20Economic%20Guide%20To%20Picking%20A%20College%20Major%20%7C%20FiveThirtyEight&parentUrl=https%3A%2F%2Ffivethirtyeight.com%2Ffeatures%2Fthe-economic-guide-to-picking-a-college-major%2F%22"

df_majors <- url %>%
  read_html() %>%
  html_nodes(css = "#example") %>%
  html_table(fill = TRUE)
```

Clean up the table
```{r clean-table}
#convert to data frame
df_majors <- data.frame(df_majors[[1]])

#change column names to first row values
colnames(df_majors) <- as.character(unlist(df_majors[1,]))

#delete first row
df_majors <- df_majors[-1, ]
```

Identify majors with "DATA" or "STATISTICS" using regex
```{r identify-majors}
library(stringr)

#convert to majors to vector of strings
majors <- as.vector(df_majors[,2])

#extract majors that contains "DATA" or "STATISTICS"
datastat_majors <- str_subset(majors, ".*(Data|Statistics).*")

datastat_majors
```
## #2
```{r convert}
#input
original <- c('[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange" 
          [5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry" 
          [9] "elderberry"   "lime"         "lychee"       "mulberry"    
          [13] "olive"        "salal berry"')

#extract words from string
extracted <- str_extract_all(original, '(?<=")\\w+\\s?\\w+?(?=")')

#generate new string of r script
newstring <- str_c(as.character(extracted), sep = ",")

message(newstring)

```
## #3
```{r subsett}
test <- c("aaa", "baab", "baba", "babcb", "abcdcba" )
#1
str_subset(test, "(.)\\1\\1")
#2
str_subset(test, "(.)(.)\\2\\1")
#3
str_subset(test, "(..)\\1")
#4
str_subset(test, "(.).\\1.\\1")
#5
str_subset(test, "(.)(.)(.).*\\3\\2\\1")
```
1. One character repeating three times in a row.

2. Two characters in the first two places which are then reversed in places 3 & 4. 

3. Two characters which are then repeated once.

4. One character, another by any character, the first character, any other character, then the first character again.

5. Three characters, zero or more characters, the three first characters but reversed.


## #4

1. "^(.).*\\1$"

2. "([A-Za-z][A-Za-z]).*\\1"

3. "([a-z]).*\\1.*\\1"

```{r test}
test <-c("dread", "church", "eleven")

#1
str_subset(test, "^(.).*\\1$")
#2
str_subset(test, "([A-Za-z][A-Za-z]).*\\1")
#3
str_subset(test, "([a-z]).*\\1.*\\1")
```

