---
title: "Assignment 5: Working with XML and JSON in R"
author: "Jordan Tapke"
date: "10/8/2020"
output:
  html_document: 
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: true
---
## Description
The assignment required saving information on three different books into a "hand created" HTML, XML, and JSON files. Once files were created, pull the information into an R data frame.

## Overview
I created the files "by hand" in Windows notebook. For the HTML book data, I saved the data into an HTML table. This proved rather easy to import into an R data frame. For the XML book data, I saved each book within a <book> element which allowed the data to be easily pulled into a data frame via the xmlToDataFrame function from XML package. For the JSON book data, I saved book objects in an array.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r, message=FALSE}
library(tidyverse)
library(XML)
library(textreadr)
library(RJSONIO)
library(rvest)
library(httr)
library(knitr)
```


## HTML
```{r, results='asis'}
html_df <- read_html(url("https://raw.githubusercontent.com/jtapke/School-Projects/master/Assignment%205/books.html")) %>%
  html_node("table") %>%
  html_table()
kable(html_df)
```

## XML
```{r, results='asis'}
url <- "https://raw.githubusercontent.com/jtapke/School-Projects/master/Assignment%205/books.xml"
res <- GET(url)
data <- xmlParse(content(res, "text"))
xml_df <- xmlToDataFrame(data)
kable(xml_df)
```

## JSON
```{r, results='asis'}
json_file <- fromJSON("https://raw.githubusercontent.com/jtapke/School-Projects/master/Assignment%205/books.json", flatten = T)[[1]]
json_df <- enframe(unlist(json_file))
json_df <- json_df %>%
  group_by(name) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = name, values_from = value) %>%
  select(-row)
kable(json_df)
```

## Comparison
All of the data frames are practically identically. The only difference was the XML data frame: the column names "Author(s)" & "Price(USD)" had to become "Authors" & "Price". This is because XML naming rules do not allow parentheses in elements.

## Conclusion
The only trouble with importing the data from these web formats is that the author(s) column is not considered tidy because multiple authors were separated by a comma. However, if this was made into "long" format it would take up considerable more memory which is not great for data that is meant to be downloaded all of the time.
